---
title: "Pre-poll Donations Breakdown"
author: "Electoral Reform Society"
date: "`r Sys.Date()`"
output:
  html_document:
    theme: "flatly"
    toc: true
    toc_float: true
    highlight: "haddock"
    css: "styles.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This page summarises the donations reported to the Electoral Commission by central party units during the pre-poll period leading up to the 2024 GE. These are also compared to donations made in 2019 by the same period. Currently, only week 1, 2 and 3 have been released for the 2024 GE. This page will be updated as more data is released.

```{r include=FALSE}

library(tidyverse)
library(knitr)
library(kableExtra)
library(plotly)
library(scales)

```

```{r functions, include=FALSE}

# remove blank spaces between £ sign and numbers for better table formatting

remove_spaces <- function(df) {
  df %>%
    mutate(across(everything(), ~ gsub("£\\s+", "£", .)))
}

# change 0's to blank cells

replace_pound_zero <- function(df) {
  df %>%
    mutate(across(everything(), ~replace(., . == "£0", "")))
}

```

### Totals: 2024 pre-poll week one - three

All donations made to the Co-operative party were combined with the Labour party, and public funds were removed. Scroll to see full tables.

```{r processing data, include=FALSE}

pp2024 <- read.csv("data/prepoll1_2024.csv")
pp2024_2 <- read.csv("data/prepoll2_2024.csv")
pp2024_3 <- read.csv("data/prepoll3_2024.csv")

#combine

pp2024 <- rbind(pp2024, pp2024_2, pp2024_3)

#removing public fund money

pp2024 <- pp2024 %>%
  filter(!DonorStatus == "Public Fund")

#combining the Labour and Cooperative party

pp2024 <- pp2024 %>%
  mutate(RegulatedEntityName = ifelse(
    RegulatedEntityName =="Co-operative Party", "Labour Party", RegulatedEntityName
  ))

# Remove £ sign and convert to numeric
pp2024$Value <- as.numeric(gsub("[£,\"]", "", pp2024$Value))

#keep only relevant columns

pp2024 <- pp2024 %>%
  select(RegulatedEntityName, ReportingPeriodName, ReceivedDate, Value, DonorStatus, DonorName)

# changing names of parties

pp2024$RegulatedEntityName <- gsub("Conservative and Unionist Party", "Conservative Party", pp2024$RegulatedEntityName)
pp2024$RegulatedEntityName <- gsub("Scottish National Party \\(SNP\\)", "SNP", pp2024$RegulatedEntityName)

#creating factors

pp2024$RegulatedEntityName <- factor(pp2024$RegulatedEntityName, levels = c("Conservative Party", "Labour Party", "Liberal Democrats", "Reform UK", "SNP", "Green Party", "Climate Party", "Workers Party of Britain", "Social Democratic Party"))
 
pp2024$DonorStatus <- factor(pp2024$DonorStatus, levels = c("Individual", "Company", "Trade Union", "Unincorporated Association", "Friendly Society"))

pp2024$ReportingPeriodName <- factor(pp2024$ReportingPeriodName, levels = c("Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024", "Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024", "Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024"))

```

```{r include=FALSE}

# Define all possible levels for DonorStatus and ReportingPeriodName
donor_statuses <- c('Individual', 'Company', 'Trade Union', 'Unincorporated Association', 'Friendly Society')
reporting_periods <- c('Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024', 'Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024', 'Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024')

# Create a complete grid of all combinations
all_combinations <- expand.grid(
  RegulatedEntityName = unique(pp2024$RegulatedEntityName),
  ReportingPeriodName = reporting_periods,
  DonorStatus = donor_statuses,
  stringsAsFactors = FALSE
)

# Merge with the original data to ensure all combinations are present
pp2024_complete <- all_combinations %>%
  left_join(pp2024, by = c("RegulatedEntityName", "ReportingPeriodName", "DonorStatus")) %>%
  mutate(Value = ifelse(is.na(Value), 0, Value))

# creating a table split by week and sum received by each party

pp_complete <- pp2024_complete %>%
  group_by(RegulatedEntityName, ReportingPeriodName, DonorStatus) %>%
  summarise(Total = sum(Value), .groups = 'drop') %>%
  unite("DonorStatus_ReportingPeriodName", DonorStatus, ReportingPeriodName) %>%
  pivot_wider(names_from = DonorStatus_ReportingPeriodName, values_from = Total, values_fill = list(Total = 0))

# redo column order

pp_complete <- pp_complete %>%
  select(RegulatedEntityName, `Individual_Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024`, `Individual_Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024`, `Individual_Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024`, `Company_Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024`, `Company_Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024`, `Company_Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024`, `Trade Union_Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024`, `Trade Union_Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024`, `Trade Union_Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024`, `Unincorporated Association_Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024`, `Unincorporated Association_Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024`, `Unincorporated Association_Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024`, `Friendly Society_Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024`, `Friendly Society_Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024`, `Friendly Society_Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024`)

```

```{r echo=FALSE}

# Format numbers as whole numbers with £ sign and remove decimal places
pp_complete <- pp_complete %>%
  mutate(across(-RegulatedEntityName, ~paste0("£", format(round(., 0), big.mark = ",", scientific = FALSE))))

# Compute row totals
pp_complete <- pp_complete %>%
  mutate(Total = rowSums(across(where(is.character), ~ as.numeric(gsub("[£,]", "", .))), na.rm = TRUE)) %>%
  mutate(Total = sprintf("£%s", format(Total, big.mark = ",", scientific = FALSE)))

# Compute column totals
column_totals <- pp_complete %>%
  summarise(across(where(is.character), ~ sum(as.numeric(gsub("[£,]", "", .)), na.rm = TRUE))) %>%
  mutate(RegulatedEntityName = "Total") %>%
  mutate(across(where(is.numeric), ~ sprintf("£%s", format(.x, big.mark = ",", scientific = FALSE))))

# Bind column totals to the original table
pp_complete <- bind_rows(pp_complete, column_totals)

# Convert the "Total" column to character to match other columns
pp_complete <- pp_complete %>%
  mutate(across(where(is.numeric), as.character))

# Remove blank spaces in column names
pp_complete <- remove_spaces(pp_complete)
pp_complete <- replace_pound_zero(pp_complete)

# Ensure the number of column names matches the actual number of columns
col_names <- c("Party", rep(c("W1", "W2", "W3"), times = 5), "")

# Summary table
pp_complete %>%
  kable("html", col.names = col_names, align = 'c') %>%
  add_header_above(c(" " = 1, "Individual" = 3, "Company" = 3, "Trade Union" = 3, "Unincorporated Association" = 3, "Friendly Society" = 3, "Total" = 1)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  row_spec(nrow(pp_complete), background = "black", color = "white") %>%
  column_spec(ncol(pp_complete), background = "black", color = "white")

```

#### 2024 totals: combined

```{r include=FALSE}

#breaking down by party and donor status

# Create a complete data frame with all combinations of Party Name and Donor Status

total_prepoll <- expand.grid(RegulatedEntityName = 
                            levels(pp2024$RegulatedEntityName), 
                         DonorStatus = 
                           unique(pp2024$DonorStatus)
                         )

# merge the values with each combo

prepoll_table <- total_prepoll %>%
  left_join(pp2024 %>%
              group_by(RegulatedEntityName, DonorStatus) %>%
            summarise(TotalValue = sum(Value, na.rm = TRUE), 
                      .groups = 'drop'),
            by = c("RegulatedEntityName", "DonorStatus"))

prepoll_table$TotalValue[is.na(prepoll_table$TotalValue)] <- 0

#Convert value to money format

prepoll_table$TotalValue <- round(prepoll_table$TotalValue)
prepoll_table$TotalValue <- sprintf("£%s", format(prepoll_table$TotalValue, big.mark = ",", scientific = FALSE))

#convert to a wide format

prepoll_table <- prepoll_table %>%
  pivot_wider(names_from = DonorStatus,
              values_from = TotalValue)

```

```{r echo=FALSE}

# Compute row totals
prepoll_table <- prepoll_table %>%
  mutate(Total = rowSums(across(where(is.character), ~ as.numeric(gsub("[£,]", "", .))), na.rm = TRUE)) %>%
  mutate(Total = sprintf("£%s", format(Total, big.mark = ",", scientific = FALSE)))

# Compute column totals
column_totals <- prepoll_table %>%
  summarise(across(where(is.character), ~ sum(as.numeric(gsub("[£,]", "", .)), na.rm = TRUE))) %>%
  mutate(RegulatedEntityName = "Total") %>%
  mutate(across(where(is.numeric), ~ sprintf("£%s", format(.x, big.mark = ",", scientific = FALSE))))

# Bind column totals to the original table
prepoll_table <- bind_rows(prepoll_table, column_totals)

# Convert the "Total" column to character to match other columns
prepoll_table <- prepoll_table %>%
  mutate(across(where(is.numeric), as.character))

# Replace 'unincorporated association' with 'unincorporated<br>association' in the column names
names(prepoll_table) <- gsub("Unincorporated Association", "Unincorporated<br>Association", names(prepoll_table))

# remove white space and replace £0

prepoll_table <- remove_spaces(prepoll_table)
prepoll_table <- replace_pound_zero(prepoll_table)

# Remove the first column title
names(prepoll_table)[1] <- ""

# Use kable to create a nicely formatted table
kable(prepoll_table, format = "html", escape = FALSE, table.attr = "style='width:100%;'", align = c("l", rep("c", ncol(prepoll_table) - 1))) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(nrow(prepoll_table), bold = TRUE, color = "white", background = "black") %>%
  add_header_above(c(" " = 1, "Donor Status" = ncol(prepoll_table) - 1), bold = TRUE) 

```

### Comparisons: 2019 and 2024 GE 

The table below summarises the total amounts received by each party, split by pre-poll donation week. The 2019 comparisons are limited to week one to three to be directly comparable.

#### Week one to three

The table below summarises totals for week one to three, split by 2019 and 2024. Interestingly, Labour received a higher total donation amount in week 2 of the 2019 campaign, however when looking at donor type, it can be seem this was massively boosted by Trade Union donations - individual and company donations remained low for Labour during week 2 of the 2019 campaign.

```{r include=FALSE}

pp2019 <- read.csv("data/alldonations_2019.csv")
pp2024 <- read.csv("data/prepoll1_2024.csv")
pp2024_2 <- read.csv("data/prepoll2_2024.csv")
pp2024_3 <- read.csv("data/prepoll3_2024.csv")

# filtering 2019 pre-poll week 1, 2 and 3

pp2019 <- pp2019 %>%
  filter(ReportingPeriodName == "Pre-Poll 1 - Party(06/11/19 - 12/11/19) UKPGE 2019"|
           ReportingPeriodName == "Pre-Poll 2 - Party(13/11/19 - 19/11/19) UKPGE 2019"|
           ReportingPeriodName == "Pre-Poll 3 - Party(20/11/19 - 26/11/19) UKPGE 2019")

#bind the tables together for comparisons

pptotal <- rbind(pp2019, pp2024, pp2024_2, pp2024_3)

#combining the Labour and Cooperative party

pptotal <- pptotal %>%
  mutate(RegulatedEntityName = ifelse(
    RegulatedEntityName =="Co-operative Party", "Labour Party", RegulatedEntityName
  ))

# removing public funds

pptotal <- pptotal %>%
  filter(!DonorStatus == "Public Fund")

#keep only relevant columns

pptotal <- pptotal %>%
  select(RegulatedEntityName, ReportingPeriodName, ReceivedDate, Value, DonorStatus, DonorName)

# Remove £ sign and convert to numeric 
pptotal$Value <- as.numeric(gsub("[£,\"]", "", pptotal$Value))

# converting names

pptotal$RegulatedEntityName <- gsub("Conservative and Unionist Party", "Conservative Party", pptotal$RegulatedEntityName)
pptotal$RegulatedEntityName <- gsub("Scottish National Party \\(SNP\\)", "SNP", pptotal$RegulatedEntityName)
pptotal$RegulatedEntityName <- gsub("Alliance - Alliance Party of Northern Ireland", "Alliance", pptotal$RegulatedEntityName)
pptotal$RegulatedEntityName <- gsub("Plaid Cymru - The Party of Wales", "Plaid Cymru", pptotal$RegulatedEntityName)

```

```{r include=FALSE}

# compare total donated to each party 

#creating factors

pptotal$RegulatedEntityName <- factor(pptotal$RegulatedEntityName, levels = c("Conservative Party", "Labour Party", "Liberal Democrats", "Reform UK", "SNP", "Green Party", "Plaid Cymru", "Climate Party", "Social Democratic Party", "Workers Party of Britain", "Alliance"))

pptotal$DonorStatus <- factor(pptotal$DonorStatus, levels = c("Individual", "Company", "Trade Union", "Unincorporated Association", "Friendly Society", "Limited Liability Partnership"))

pptotal$ReportingPeriodName <- factor(pptotal$ReportingPeriodName, levels = c("Pre-Poll 1 - Party(06/11/19 - 12/11/19) UKPGE 2019", "Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024", "Pre-Poll 2 - Party(13/11/19 - 19/11/19) UKPGE 2019", "Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024", "Pre-Poll 3 - Party(20/11/19 - 26/11/19) UKPGE 2019", "Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024"))

# Create a complete data frame with all combinations of Party Name and pre-poll period

pp_comparison <- expand.grid(RegulatedEntityName = 
                            levels(pptotal$RegulatedEntityName), 
                         ReportingPeriodName = 
                           unique(pptotal$ReportingPeriodName)
                         )

# merge the values with each combo

prepoll_comparison <- pp_comparison %>%
  left_join(pptotal %>%
              group_by(RegulatedEntityName, ReportingPeriodName) %>%
            summarise(TotalValue = sum(Value, na.rm = TRUE), 
                      .groups = 'drop'),
            by = c("RegulatedEntityName", "ReportingPeriodName"))

prepoll_comparison$TotalValue[is.na(prepoll_comparison$TotalValue)] <- 0

#Convert value to money format

prepoll_comparison$TotalValue <- round(prepoll_comparison$TotalValue)
prepoll_comparison$TotalValue <- sprintf("£%s", format(prepoll_comparison$TotalValue, big.mark = ",", scientific = FALSE))

#convert to a wide format

prepoll_comparison <- prepoll_comparison %>%
  pivot_wider(names_from = ReportingPeriodName,
              values_from = TotalValue)

#rename column names

prepoll_comparison <- prepoll_comparison %>%
  rename(`Week 1 2019` = `Pre-Poll 1 - Party(06/11/19 - 12/11/19) UKPGE 2019`,
         `Week 1 2024` = `Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024`,
         `Week 2 2019` = `Pre-Poll 2 - Party(13/11/19 - 19/11/19) UKPGE 2019`,
         `Week 2 2024` = `Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024`,
         `Week 3 2019` = `Pre-Poll 3 - Party(20/11/19 - 26/11/19) UKPGE 2019`,
         `Week 3 2024`= `Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024`)

# reorder columns

prepoll_comparison <- prepoll_comparison %>%
  select(RegulatedEntityName, `Week 1 2019`, `Week 1 2024`, `Week 2 2019`, `Week 2 2024`, `Week 3 2019`, `Week 3 2024`)

# Calculate column totals
column_totals <- prepoll_comparison %>%
  summarise(across(where(is.character), ~ sprintf("£%s", format(sum(as.numeric(gsub("[£,]", "", .))), big.mark = ",", scientific = FALSE)))) %>%
  mutate(RegulatedEntityName = "Total")

# Bind the totals row to the original table
prepoll_comparison <- bind_rows(prepoll_comparison, column_totals)

# remove spaces and replace £0 with blanks

prepoll_comparison <- remove_spaces(prepoll_comparison)
prepoll_comparison <- replace_pound_zero(prepoll_comparison)

# format table

totals_comparison <- kable(prepoll_comparison, col.names = c("", "2019", "2024", "2019", "2024", "2019", "2024"), align = 'c') %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Week One" = 2, "Week Two" = 2, "Week 3" = 2)) %>%
  row_spec(nrow(prepoll_comparison), bold = TRUE, color = "white", background = "black") 

```

```{r, results='asis', echo=FALSE}

totals_comparison

```

The graph below shows the same trends for the main parties that received donations: Conservatives, Labour, Lib Dems, Reform, Greens, Plaid and SNP. 

```{r echo=FALSE, fig.align='center'}

# graph to show the same thing with more clarity

pptotal <- pptotal %>%
  mutate(PrePollWeek = case_when(
    ReportingPeriodName == "Pre-Poll 1 - Party(06/11/19 - 12/11/19) UKPGE 2019" ~ "week one",
    ReportingPeriodName == "Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024" ~ "week one",
    ReportingPeriodName == "Pre-Poll 2 - Party(13/11/19 - 19/11/19) UKPGE 2019" ~ "week two",
    ReportingPeriodName == "Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024" ~ "week two",
    ReportingPeriodName == "Pre-Poll 3 - Party(20/11/19 - 26/11/19) UKPGE 2019" ~ "week three",
    ReportingPeriodName == "Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024" ~ "week three",
    TRUE ~ ReportingPeriodName
  ))

# merge the reporting period names into 2019 and 2024 totals

pptotal <- pptotal %>%
  mutate(Year = case_when(
    ReportingPeriodName == "Pre-Poll 1 - Party(06/11/19 - 12/11/19) UKPGE 2019" ~ "2019",
    ReportingPeriodName == "Pre-Poll 2 - Party(13/11/19 - 19/11/19) UKPGE 2019" ~ "2019",
    ReportingPeriodName == "Pre-Poll 1 - Party(30/05/24 - 05/06/24) UKPGE 2024" ~ "2024",
    ReportingPeriodName == "Pre-Poll 2 - Party(06/06/24 - 12/06/24) UKPGE 2024" ~ "2024",
    ReportingPeriodName == "Pre-Poll 3 - Party(20/11/19 - 26/11/19) UKPGE 2019" ~ "2019",
    ReportingPeriodName == "Pre-Poll 3 - Party(13/06/24 - 19/06/24) UKPGE 2024" ~ "2024",
    TRUE ~ ReportingPeriodName
  ))

# setting factor levels for week

pptotal$PrePollWeek <- factor(pptotal$PrePollWeek, level = c("week one", "week two", "week three"))

# filter out for main parties

ppmain <- pptotal %>%
  filter(RegulatedEntityName == "Conservative Party"|
           RegulatedEntityName == "Labour Party"|
           RegulatedEntityName == "Liberal Democrats"|
           RegulatedEntityName == "Reform UK"|
           RegulatedEntityName == "Green Party"|
           RegulatedEntityName == "Plaid Cymru"|
           RegulatedEntityName == "SNP")

# creating a line graph that has week along the x-axis but splits into facets by year

# Aggregate the data to sum donations by party, period, and donor status
line_party <- ppmain %>%
  group_by(RegulatedEntityName, PrePollWeek, Year) %>%
  summarise(TotalValue = sum(Value), .groups = 'drop')

# Hover text customised

line_party <- line_party %>%
  mutate(HoverText = paste0("Total Value: £", format(round(TotalValue, 0), big.mark = ",", scientific = FALSE), "<br>",
                            "Party: ", RegulatedEntityName, "<br>",
                            "Pre-poll period: ", PrePollWeek))

# customised colours

party_colors <- c(
  "Conservative Party" = "#000080",  # Navy Blue
  "Labour Party" = "#FF0000",        # Red
  "Liberal Democrats" = "#FFD700",   # Dark Yellow
  "Reform UK" = "#ADD8E6",           # Light Blue
  "Green Party" = "#006400",         # Dark Green
  "Plaid Cymru" = "#90EE90",         # Light Green
  "SNP" = "orange"                  # Orange
)

# Create the line chart

p <- ggplot(line_party, aes(x = PrePollWeek, y = TotalValue, color = RegulatedEntityName, text = HoverText, group = RegulatedEntityName)) +
  geom_line(linewidth = 0.5) +
  geom_point(size = 2) +
  scale_y_continuous(labels = dollar_format(prefix = "£", big.mark = ",")) +
  labs(x = "", y = "", color = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_color_manual(values = party_colors) +
  facet_wrap(~ Year, scales = "free_x", nrow = 1) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))

# Convert to interactive plot with plotly
maintrends_interactive <- ggplotly(p, tooltip = "text")

# show

maintrends_interactive

```

#### Split by donor type

The table below summarises the total amounts recieved in week 1, 2 and 3 of the campaigns, split by donor type (scroll left to see all columns). 

```{r echo=FALSE, warning=FALSE}

# create summary table

year_summary <- pptotal %>%
  select(RegulatedEntityName, Year, DonorStatus, Value) %>%
  group_by(RegulatedEntityName, DonorStatus, Year) %>%
  summarise(TotalValue = sum(Value), .groups = 'drop') %>%
  complete(RegulatedEntityName, DonorStatus, Year, fill = list(TotalValue = 0))

# Convert value to money format
year_summary$TotalValue <- round(year_summary$TotalValue)
year_summary$TotalValue <- sprintf("£%s", format(year_summary$TotalValue, big.mark = ",", scientific = FALSE))

# create wide format 

year_summary_wide <- year_summary %>%
  pivot_wider(
    names_from = c(DonorStatus, Year),
    values_from = TotalValue,
    names_sep = "_"
  )

# Calculate column totals

column_total <- year_summary_wide %>%
  summarise(across(everything(), ~ sum(as.numeric(gsub("[£,]", "", .)),
                                       na.rm = TRUE))) %>%
  mutate(RegulatedEntityName = "Total") %>%
  mutate(across(-RegulatedEntityName, ~ sprintf("£%s", format(.x, big.mark = ",", scientific = FALSE))))

# remove blank spaces function

year_summary_wide <- remove_spaces(year_summary_wide)

# use replace_pound_zero function to make a neater table

year_summary_wide <- replace_pound_zero(year_summary_wide)

# bind the new row with original dataframe

year_summary_wide <- rbind(year_summary_wide, column_total)

```

```{r echo=FALSE}

# Replace 'unincorporated association' with 'unincorporated<br>association' in the column names
names(year_summary_wide) <- gsub("Unincorporated Association", "Unincorporated<br>Association", names(year_summary_wide))

# Remove the first column title
names(year_summary_wide)[1] <- ""

#create kable table

kable(year_summary_wide, col.names = c("", "2019", "2024", "2019", "2024", "2019", "2024", "2019", "2024", "2019", "2024", "2019", "2024")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = FALSE) %>%
  add_header_above(c(" " = 1, "Individual" = 2, "Company" = 2, "Trade Union" = 2, "Unincorporated Association" = 2, "Friendly Society" = 2, "Limited Liability Partnership" = 2)) %>%
  row_spec(nrow(year_summary_wide), bold = TRUE, color = "white", background = "black") %>%
  row_spec(0, align = "center")


```

### Biggest party comparisons

Below sums up the totals for week one to three received in 2019 and 2024 by the key parties. This is to get a better sense of total amounts compared to 2019 - it is interesting that there was still more money going into the Conservatives relative to how much Labour are getting now.

This may be related to the increase in reporting threshold; in 2019 any donation over £7,500 needed to be reported, the Elections Act of 2022 raised this threshold to £11,180. Rather than less money being accepted, it may be less is required to be reported. This is speculation.

```{r echo=FALSE, fig.align='center'}

pptotal <- pptotal %>%
  filter(RegulatedEntityName == "Conservative Party"|
         RegulatedEntityName == "Labour Party"|
           RegulatedEntityName == "Liberal Democrats"|
           RegulatedEntityName == "Reform UK")

# this leave 350 donations

# Aggregate the data to sum donations by party, period, and donor status
pptotal_agg <- pptotal %>%
  group_by(RegulatedEntityName, Year, DonorStatus) %>%
  summarise(TotalValue = sum(Value), .groups = 'drop')

# Hover text customised

pptotal_agg <- pptotal_agg %>%
  mutate(HoverText = paste0("Total Value: £", format(round(TotalValue, 0), big.mark = ",", scientific = FALSE), "<br>",
                            "Donor Status: ", DonorStatus, "<br>",
                            "Party: ", RegulatedEntityName, "<br>",
                            "Year: ", Year))

# Create the bar chart
p <- ggplot(pptotal_agg, aes(x = RegulatedEntityName, y = TotalValue, fill = DonorStatus, text = HoverText)) +
  geom_bar(stat = "identity") +
  labs(x = "", y = "", fill = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        legend.position = "bottom",
        legend.justification = "center") +
  scale_y_continuous(labels = scales::label_dollar(prefix = "£", big.mark = ",")) +
  scale_fill_brewer(palette = "Set3") +
  facet_wrap(~ Year, scales = "free_x", nrow = 1) +
  theme(strip.background = element_blank(),
        strip.text = element_text(face = "bold"))

# Convert to interactive plot with plotly
totals_interactive <- ggplotly(p, tooltip = "text")

# show

totals_interactive

```